### 2014-07-03 Paper1/paper1.tex
Updates from home

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test the approach using human reaching data. We first show how to mathematically invert typical feedforward control approaches, and then examine how this reveals the strengths, limitations, and underlying assumptions. We found that sensitivity to parameter inaccuracies was on the order of microns and hence much lower than the inaccuracies seen when humans reach to visual targets. Further, most of the sensitivity resulted from uncertainty with regard to the joint stiffness. We also tested the robustness of this approach on human point-to-point reaching movements disturbed by force pulses. As expected, the intended trajectory showed no change from undisturbed reaching for 300 ms after the disturbance onset. Beyond this point, we detected change in intent even when the intention was near the target.  Unlike well-known $\lambda$ models, this intent was determined by considering both feedforward (predictive) and feedback controls. Knowing intent alongside action allows real-time determination and manipulation of error, even without a specified task, which makes it possible to directly inspect a neurally-impaired patient's intent for deficits. This approach may also be scaled for determining intent during human-machine interaction and in complex systems like team sports, crowds, or swarms.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}
In our everyday struggles with bumpy roads and bustling crowds we often make movements that do not match what we had planned. Here we identify a computational modeling method that measures actions and allows us to determine what the person meant to to do. We first demonstrate this method's good performance even when there is uncertainty in the parameters of the model. Next, we evaluate this on human reaching motions to targets, and show that while the arm is bumped by a disturbance, the intention remains straight. Approximately 300 milliseconds later, the intention changes in a corrective action that counteracts the disturbance. This type of computation can be broadly useful in canceling the effects of disturbances such as turbulence while typing or piloting, and can estimate intent in many other dynamic systems from molecules to crowds of people. It is also a new tool for understanding and treating impaired movement plans in neural injury.

@@

It is important to distinguish our use of the term intent from other groups that consider \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}. In typical motor control studies, subjects are motivated to complete the experiment in a timely fashion and are usually explicitly provided targets to be reached. Classification of intent is prevalent in both lower \cite{strausser2011development, hargrove2013robotic} and upper limb\cite{englehart2003robust, young2012improving} prosthetics where hybrid control algorithms select from among a set of actions (walking/standing/flexion/extension/etc). While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves. Particularly of interest is the intended course of action, even when the actual movement is disturbed and hence no longer matches the intent.

Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles. Springs produce a force according to their impedance and stretch. By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point. Supporters of the "$\lambda$ model"\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement. Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable. Therefore, it could not well-represent the intent of a simple reaching movement. We propose a revised method---intent extraction---that deduces intended trajectory, even in the face of disturbances. This method describes a class of filters that map a disturbance, process, actuator, and realized trajectory to an intended trajectory. In human reaching, the process (a dynamic model of the arm), the actuator (viscoelastic muscles), and the realized trajectory are all readily accessible via measurement and modeling. The mathematics underlying intent determination are sufficiently general to be extensible to whole-body movement or even more abstract, psychological processes.

Here we develop and analyze a method to extract intent from disturbed movement. This method then allows us to observe if and how intent changes following disturbance. We employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction. We then explore sensitivity and failure modes using standard model analysis techniques. We also demonstrate the accuracy and utility of this on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances. We hypothesize that this combination allows intent extraction with less uncertainty than the uncertainty naturally inherent in point-to-point reaching. We show that certain factors and measurements are more sensitive than others, and demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed.

@@

\subsection*{Intent Extraction: Deducing the Desired Trajectory}
We first demonstrate the \textit{intent extraction} approach in human motor control, but we later show that the process is applicable to any controlled process with an invertible highest order plant term. The process begins by presuming a model of the controller. Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and linearizes the system through cancellation. Additionally, linear feedback rejects position and velocity error. This model was chosen to help illuminate the approach, but its feedback component lacks the complexity and accuracy of more recent approaches. Nevertheless later we will discuss how this approach can be used on more complex models without loss of generality. The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
\end{equation}
Where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, $G$ contains both Coriolis and centripetal effects, and $E$ is any externally-applied torque. The motion behavior changes with the addition of feedforward and/or feedback controllers:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
\end{equation}
Some terms have hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) to indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{kawato1999internal}. This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm in the torque balance. If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero. $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$. This $q_d$ signifies the unknown desired trajectory that we seek to discover.

@@

While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model. Instead of a dynamic model, it could be a lookup table. For instance, a light switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism. In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance. As long as some bidirectional relationship (even statistical) exists between state and outcome, intended outcome can be determined. This allows determination of intention in many situations of interest, even where the process is irreducibly complex.

\subsection*{Generation of synthetic data using simple dynamic simulation}
While the derivation above is general, testing its application to human reaching requires choices for plant and actuator models. Appropriate plant modeling is well-understand: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms. Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia. For the actuator, we choose the model of Burdet et al. \cite{burdet2006stability}. A scalar multiple of their modeled feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation} while leaving the shape and structure of the impedance unchanged. As a consequence of choosing this model, muscle torque must be used to calculate stiffness even though it arises in part from stiffness. Like Burdet et al. \cite{burdet2006stability}, we approximated muscle torque by looking backward in time a small amount (500 microseconds). This approximation is not necessary during intent extraction, which leads to a small (sub-micrometer mean unsigned error) discrepancy. Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration.

\subsection*{Experimental Design}
Simulated reaches were 15 centimeters long, beginning 38 centimeters out from and 5.7 centimeters left of the right shoulder then proceeding right.  We used the following types of perturbing forces for each combination of distance and direction: 
pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion. The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.

\subsection*{Error Metric}
Our error metric used deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the nominal trajectory across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15 centimeter reach.  Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis. Mean and maximum signed error relative to the direction of disturbance, also called perpendicular deviation, were used to measure reaching accuracy in order to elucidate direction of any corrections.

\subsection*{Indices of variance-based sensitivity}
Sobol-distributed matrices were generated using MATLAB's sobolset() function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density. Distributions were clamped to the range of $\pm3$ standard deviations in order to limit parameters to a realistic range. Combinations and calculations were made as prescribed by Saltelli et al. \cite{saltelli2010variance} to arrive at direct and total sensitivity indices.

@@

\subsection*{Apparatus}
A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any friction or mass. The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.  Visual feedback was performed at 60 Hz using OpenGL. Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.

\subsection*{Protocol}
Subjects made 730 reaches in total, along a line parallel to their coronal plane and approximately 45 centimeters from their shoulder. Reaches were either 15 or 30 centimeters long, starting and ending at one of three points spaced 15 centimeters apart on the line. To prevent any learning effect, forces were presented intermittently with random frequency, but never less than 5 reaches apart. Forces were chosen pseudorandomly such that each type, direction, and distance combination mentioned above was presented 5 times.

\subsection*{Statistical Analysis}
The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons among subjects and trajectory sources using ANOVA with Tukey-Kramer post-\textit{hoc} test for group differences.

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Model Robustness}
While some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input. Sensitivity analysis revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.  Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (Figure 1) were extracted in the presence of 220,000 variations upon the parameters (Table 1) according to the methods of Saltelli et al. \cite{saltelli2010variance}. Expected variance due to parameter uncertainty was $0.86  \mu m^2$ for pulse forces and $1.5 \mu m^2$ for filtered white forces. Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 \mu m^2$. These variances all fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter).

\subsection*{Parameter Sensitivity}
Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to misestimation of stiffness and changes in shoulder position from trial to trial. Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all other parameters were accurate,  and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty. Note that this sensitivity analysis does not provide information on whether models or parameter estimates are accurate, only how sensitive the model would be if they were inaccurate.

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3). In both pulse timing conditions, the intended trajectory remained straight towards the target, supporting the approach.  When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred late in reaching, intent continued onto the target but then corrected anyways (Figure 3 lower plots). We observed no significant change of intent during force disturbance until 300 milliseconds after the onset of disturbance. Intention showed some systematic deflection toward the direction of disturbance peaking around 150 milliseconds after the onset of force disturbance, but this effect did not leave the range of error found in undisturbed movement. While the hand returns within 600 milliseconds to undisturbed levels of error, intention appears to first over-correct and is altered for as much as one second (Figure 4 shaded areas).

We argue that the approach here represents the first sound method for looking at intent following disturbance, so we have only the undisturbed intention for comparison. If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands. This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay. We test this in Figure 4.

Following disturbance, we compare deviation in the hand and intended paths against the maximum deviation of undisturbed movements in order to determine at what point in time changes in hand path or intent can no longer be explained by normal variation and inaccuracy. Subjects' force-disturbed hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement beginning within at most 135 milliseconds after the onset of disturbance. By contrast, intent did not change significantly for at least 300 milliseconds after the onset of disturbance (Table 2). This passes the test described above. While there does appear to be some systematic failure of the intent model to completely reject the disturbance, this residual disturbance does not exceed the statistical threshold for significance. An overcorrection beginning at approximately 200 milliseconds appears to account for change in intention in all subjects.

\section*{Discussion}
We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there are disturbances. Sensitivity analysis on synthetic data revealed that the method was suitable in that errors in response to parameter variations were smaller than trial-to-trial variance observed in human reaching. Significant misestimation could occur if the stiffness was inaccurately modeled. We tested this on human reaching and found that in spite of the hand being distubed by forces, the intended movement remained on-course to the target, perhaps changing between 200 and 300 milliseconds. This represents a new and accurate method for viewing intent and how it changes when faced with a force disturbance.

@@

The proposed approach could explain the discovery by Bizzi et al. \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$. By deriving the technique in a general form, we discovered that our $x_d$ should be equivalent to $\alpha$ and can be used to determine the muscle equilibrium of a $\lambda$ model. $x_d$ and $\lambda$ are only equivalent when $x_d$ is not changing.  Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of a virtual mass. This provides a striking and falsifiable prediction: reaching intention cannot jump abruptly.

Our analysis suggests that intent is slow to change if it changes at all following disturbance. While this work examines the time of onset of change, statistical testing cannot be used to detect the absence of change in an individual movement. Still, Figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control. In classic optimal control, the system is always updating the intent in response to errors \cite{todorov2002optimal}. Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}. Interestingly, disturbances that come later in the movement (Figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly ``stick to their guns'' when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target. Time of response may be dependent on the time of the disturbance in the movement. Alternatively, control might be intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent. It remains to be seen if and how change in intent might be triggered.

@@

The process model does not need to be explicitly formulated. Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from national security to athletic performance. The plant (such as the actions of a crowd in a crowded square or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above. However, by substituting some approximation or lookup table of input-outcome pairs, it is possible in a variety of applications to obtain the intent, even when disturbed. In other words, it may not be necessary to allow a bomber to succeed or an opposing team to be victorious to understand the plan behind their action.

>>>>

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test the approach using human reaching data. We first show how to mathematically invert typical feedforward control approaches, and then examine how this reveals the strengths, limitations, and underlying assumptions. We found that sensitivity to parameter inaccuracies was on the order of microns and hence much lower than the inaccuracies seen when humans reach to visual targets. Further, most of the sensitivity resulted from uncertainty with regard to the joint stiffness. We also tested the robustness of this approach on human point-to-point reaching movements disturbed by force pulses. As expected, the intended trajectory showed no change from undisturbed reaching for 300 ms after the disturbance onset. Beyond this point, we detected change in intent even when the intention was near the target.  Unlike well-known lambda ($\lambda$) models, this intent was determined by considering both feedforward (predictive) and feedback controls. Knowing intent alongside action allows real-time determination and manipulation of error, even without a specified task, which makes it possible to directly inspect a neurally-impaired patient's intent for deficits. This approach may also be scaled for determining intent during human-machine interaction and in complex systems like team sports, crowds, or swarms.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}
In our everyday struggles with bumpy roads and bustling crowds we often make movements that do not match what we had planned. Here we identify a computational modeling method that measures actions and allows us to determine what the person meant to to do. We first demonstrate this method's good performance even when there is uncertainty in the parameters of the model. Next, we evaluate this on human reaching motions to targets, and show that while the arm is bumped by a disturbance, the intention remains straight. Approximately 300 milliseconds later, the intention changes in a corrective action that counteracts the disturbance. This type of computation can be broadly useful in canceling the effects of disturbances such as turbulence while typing or piloting, and can estimate intent in many other dynamic systems from molecules to crowds of people. It is also a new tool for understanding and treating impaired movement plans following neural injury.

@@

It is important to distinguish our use of the term intent from other groups that consider \textit{motivation}\cite{mcclelland1985motives, rawolle2013relationships}, \textit{cost}\cite{todorov2002optimal, flash1985coordination}, or \textit{goal selection decisions}\cite{ziebart2010modeling}. In typical motor control studies, subjects are motivated to complete the experiment in a timely fashion and are usually explicitly provided targets to be reached. Classification of intent is prevalent in both lower \cite{strausser2011development, hargrove2013robotic} and upper limb\cite{englehart2003robust, young2012improving} prosthetics where hybrid control algorithms select from among a set of actions (walking/standing/flexion/extension/etc). While subjects may be \textit{motivated} to complete experiments with minimal effort/cost and their \textit{goal} may be to reach a target, here we use intent to describe the course of action (i.e., the trajectory of the arm) taken in service of goals and motives and not the goals nor the motives themselves. Particularly of interest is the intended course of action (i.e, the intended trajectory), even when the actual movement is disturbed and hence no longer matches the intent.

Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles. Springs produce a force according to their impedance and stretch. By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point. Supporters of the "$\lambda$ model"\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement. Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable. Therefore, it could not well-represent the intent of a simple reaching movement. We propose a revised method --- \textit{intent extraction} --- that deduces intended trajectory, even in the face of disturbances. This method describes a class of filters that map a disturbance, process, actuator, and realized trajectory to an intended trajectory. In human reaching, these initial ingredients are available: the process is a dynamic model of the arm, the actuators (muscles) are assumed to act as viscoelastic torque generators, the human interface forces are measured by a sensor, and the realized trajectory is measured.

Here we develop and analyze the mathematics underlying intent extraction, a method to determine one's intent. We focused, for simplicity in this initial study, on planar targeted arm reaching actions. First, we produce synthetic reaching data using a forward model so that we can examine the quality and uncertainty of intent extraction. We then explore sensitivity and failure modes, demonstrating the accuracy and utility of this method. We show efficacy on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances. Finally, we demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed, showing how one's intent may change following a disturbance.

% M E T H O D S

@@

\subsection*{Intent Extraction: Deducing the Desired Trajectory}
We first demonstrate the \textit{intent extraction} approach in human motor control, but we later show that the process is applicable to any controlled process with an invertible highest order plant term. The process begins by presuming a model of the controller. Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and linearizes the system through cancellation. Additionally, linear feedback rejects position and velocity error. This model was chosen to help illuminate the approach, but its feedback component lacks the complexity and accuracy of more recent approaches. Nevertheless later we will discuss how this approach can be used on more complex models without loss of generality. The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
\end{equation}
Where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, $G$ contains both Coriolis and centripetal effects, and $E$ is any externally-applied torque. The motion behavior changes with the addition of feedforward and/or feedback controllers:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
\end{equation}
Some terms have hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) to indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{shadmehr1994adaptive}. This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm in the torque balance. If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero. $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$. This $q_d$ signifies the unknown desired trajectory that we seek to discover.

@@

While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model. Instead of a dynamic model, it could be a lookup table. For instance, a switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism. In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance. As long as some bidirectional relationship (even statistical) exists between state and outcome, intended outcome can be determined. This allows determination of intent in many situations of interest, even where the process is irreducibly complex.

\subsection*{Generation of synthetic data using simple dynamic simulation}
While the derivation above is general, testing its application to human reaching requires choices for plant and actuator models. Here we show the results from our choice of Burdet and colleagues \cite{burdet2006stability}. Appropriate plant modeling is well-understood: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms. Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia. Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration \cite{flash1985coordination}. For the actuator, we choose the model of Burdet et al. \cite{burdet2006stability}. A scalar multiple of their modeled feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation} while leaving the shape and structure of the impedance unchanged. In this  model, musculoskeletal stiffness was linearly related to torque. Like Burdet et al. \cite{burdet2006stability}, we approximated muscle torque by looking backward in time 500 microseconds. Note that this backward look led to negligible (sub-micrometer mean unsigned error) discrepancies in subsequent intent estimates.

\subsection*{Experimental Design}
We chose 15 centimeters long simulated reaches of the right arm, beginning 38 centimeters out from and 5.7 centimeters left of the right shoulder then first proceeding to the right.  We used the following two types of perturbing forces for each combination of distance and direction: 
(1) pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
(2) Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion. The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10 \pi$ rad/s.

Our error metric was deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the straight-line nominal trajectory to the target. This was evaluated across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15 centimeter reach. Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis. Mean and maximum signed error relative to the direction of disturbance, also called perpendicular deviation, were used to measure reaching accuracy in order to elucidate direction of any corrections.

\subsection*{Indices of variance-based sensitivity}
To ensure a properly spaced and efficient evaluation of sensivity, we employed Sobol-distributed matrices, generated using MATLAB's \textit{sobolset()} function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density. Distributions were clamped to the range of $\pm3$ standard deviations in order to limit parameters to a realistic range. Combinations and calculations were made as prescribed by Saltelli et al. \cite{saltelli2010variance} to arrive at direct and total sensitivity indices.

@@

\subsection*{Apparatus}
A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any friction or mass. The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.  Visual feedback was performed at 60 Hz using OpenGL. Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame. Because force sensors tend to drift, we performed a linear re-zeroing procedure between each motion to assure unbiased measurements.

\subsection*{Protocol}
Subjects made 730 reaches in total, along a line parallel to their coronal plane and approximately 45 centimeters from their shoulder. Reaches were either 15 or 30 centimeters long, starting and ending at one of three points spaced 15 centimeters apart on the line. To prevent any learning effect, forces were presented intermittently with random frequency, but never less than 5 reaches apart. Forces were chosen pseudorandomly such that each type, direction, and distance combination mentioned above was presented 5 times. We used these disturbed, intermittent trials for the analysis.

\subsection*{Statistical Analysis}
The MATLAB statistics toolbox package \cite{MATLAB:2008} was used to perform comparisons among subjects and trajectory sources using ANOVA with Tukey-Kramer post-\textit{hoc} test for group differences. We used an alpha level of 0.05 for declaring statistical significance.

% R E S U L T S

% Results and Discussion can be combined.
\section*{Results}
As expected, the model was able to recover the original intended trajectory even when disturbed (Figure 1). However this idealized analysis cannot reveal any vulnerabilities to unexamined model parameters or to inaccuracy in the structure of the model itself as discussed in the next sections.

\subsection*{Noise Robustness}
While some inversion processes might be highly sensitive to noise, our process for recovering intent from action did not appear to be vulnerable. Our position sensors are very accurate (no detectable drift, RMS noise 1.1 micrometers), but the force sensors are less so. Even after correcting for drift in our force measurements, the force measurements reported during reaching can be expected to have noticeable bias and further noise on the order of that bias as reported in Table 1. Taken together, force measurement errors account for 21\% of all sensitivity. We also examined noise in sensor signals explicitly in the more comprehensive  sensitivity analysis, discussed below.

\subsection*{Parameter Sensitivity}
Variance in intended trajectories due to estimated uncertainty in model parameters was lower than the natural variation in undisturbed motions. Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (Figure 1) were extracted in the presence of 220,000 variations upon the parameters (Table 1) according to the methods of Saltelli et al. \cite{saltelli2010variance}. Expected variance due to parameter uncertainty was $0.86  \mu m^2$ for pulse forces and $1.5 \mu m^2$ for filtered white forces. Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 \mu m^2$. These variances all fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter).

Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to mis-estimation of stiffness and changes in shoulder position from trial to trial. Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all other parameters were accurate,  and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty. Note that this sensitivity analysis does not provide information on whether models or parameter estimates are accurate, only how sensitive the model would be if they were inaccurate.

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3). If intent can change following disturbance, it can only do so after some period of time due to processing and communication delays in the sensorimotor system that produces new motor commands. We hypothesized that intent should not diverge from the undisturbed intent path even if the actual hand was disturbed within this window of delay. In both pulse timing conditions, the intended trajectory remained straight towards the target, even though the actual hand was dramatically deflected, supporting our hypothesis and the approach.

These results suggested a reproducible approach for inspecting how individuals might alter their intent in response to disturbances. Actual hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement beginning within at most 135 milliseconds after the onset of disturbance. We observed no significant change of intent during force disturbance until 300 milliseconds after the onset of disturbance (Table 2). When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred later, the intent continued to the target, but interestingly, it then corrected (Figure 3 lower plots). Intention showed systematic deflections that counteracted the direction of disturbance about 200 milliseconds after the onset of the force disturbance. This counteraction did not depart from the range of undisturbed movements, however (Figure 4, green region). The movements completed with the hand returning to undisturbed levels of error within 600 milliseconds, intention differed from actual hand path for as long as one second after the onset of motion.

% D I S C U S S I O N

\section*{Discussion}
We sought to test the suitability and robustness of an algorithm that determines a person's intent during reaching motion, even if there are disturbances. Sensitivity analysis on synthetic data revealed that errors in response to parameter variations were smaller than the trial-to-trial variance commonly observed in human reaching. Mis-estimation was largest if the stiffness was inaccurately modeled. We tested this on human reaching and found that in spite of the hand being disturbed by forces, the intended movement remained on-course to the target, perhaps changing between 200 and 300 milliseconds. This represents a new and accurate method for viewing intent and how it changes when faced with a force disturbance.

@@

The proposed approach could explain the discovery by Bizzi et al. \cite{bizzi1984posture} of a ``virtual trajectory'' ($\alpha$) in deafferented monkeys that progressed smoothly from the initial to final positions and how that virtual trajectory is different from $\lambda$, the equilibrium point of muscles -- $x_d$ and $\lambda$ are equivalent only when $x_d$ is unchanging.  By deriving the technique in a general form, we discovered that our $x_d$ is equivalent to $\alpha$ and can be used to determine the muscle equilibrium of a $\lambda$ model.

Interestingly, the formulation leads to some implications on the expected behavior of this moving equilibration. Even if the muscle equilibrium jumped abruptly, this would be insufficient to move the intended trajectory abruptly due to the presence of the mass term in the intention formulation -- reaching intent cannot jump abruptly. The intent trajectories certainly follow this pattern, smoothly responding to disturbances at a latency of about 200 milliseconds. While this work examines the tendencies of the onset of change, statistical testing cannot be used to detect the absence of change in an individual movement. Still, Figure 3 shows reaches that are more compatible with intermittent control than with continuous optimal feedback control. In classic optimal control, the system is always updating intent in response to error \cite{todorov2002optimal}. Our results suggest a latent response consistent with sensory feedback latencies \cite{pruszynski2012optimal}. Interestingly, disturbances that come later in the movement (Figure 3, bottom panel) appeared to not dramatically affect some subjects' intent, suggesting that people may stubbornly persist with the same intent when they know that the stiffness and damping characteristics will provide enough restoring torque to arrive at the target. Time of response may be dependent on the time of the disturbance in the movement. Alternatively, control might be intermittent \cite{gawthrop2011intermittent} with limited opportunities to change intent. It remains to be seen if and how change in intent might be triggered.

@@

The process model does not need to be explicitly formulated. Dynamic models that strive to understand intent are wide ranging, and can include mathematical models that range in application from swarm prediction to athletic performance. The plant (such as the actions of a crowd of people or the motions of an opposing team) may be too complex to model with simple linear transformations as we present above. However, by substituting any approximation or lookup table of input-outcome tendencies, it is possible in a variety of applications to obtain the intent, even when disturbed. In other words, it may not be necessary to allow a bomber to succeed or an opposing team to be victorious to uncover the plan behind their action.
